High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-20 paper/main.tex
added some notes
working on introduction (backing up progress)
added a paragraph about information to the intro

Large-scale networks, including the human brain, may be conceptualized as
occupying one or more positions along on a continuum. At one extreme, every
node is fully independent of every other node. At the other extreme, all nodes
are fully coupled and behave identically. Each extreme optimizes key properties
of how the network functions. When every node is independent, the network is
maximally \textit{expressive}: if we define the network's ``state'' as the
total set of activity patterns across nodes, then every state is equally
reachable by a network with fully independent nodes. On the other hand, a fully
coupled network optimizes \textit{robustness}: any subset of nodes, other than
the entire network, may be removed from the network without any loss of
function or expressive power. Note that a given set of nodes might reconfigure
its connections or behaviors under different circumstances to change its
position along this continuum according to the needs at hand. Presumably, most
systems tend to occupy positions between the above extremes. We wondered: might
the human brain reconfigure itself to be more flexible or more robust according
to ongoing demands?

>>>>

Large-scale networks, including the human brain, may be conceptualized as
occupying one or more positions along on a continuum. At one extreme, every
node is fully independent of every other node. At the other extreme, all nodes
are fully coupled and behave identically. Each extreme optimizes key properties
of how the network functions. When every node is independent, the network is
maximally \textit{expressive}: if we define the network's ``state'' as the
total set of activity patterns across nodes, then every state is equally
reachable by a network with fully independent nodes. On the other hand, a fully
coupled network optimizes \textit{robustness}: any subset of nodes, other than
the entire network, may be removed from the network without any loss of
function or expressive power. Note that a given set of nodes might reconfigure
its connections or behaviors under different circumstances to change its
position along this continuum according to the needs at hand. Presumably, most
systems tend to occupy positions between these extremes. We wondered: might
the human brain reconfigure itself to be more flexible or more robust according
to ongoing demands?

% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.

Closely related to the above notions of expressiveness versus robustness are
measures of how much \textit{information}~\citep{Shan48} is contained in a
given signal or pattern. Formally, information is defined as the amount of
uncertainty (i.e., entropy) in a given variables' outcomes, measured in
\textit{bits}, or the optimal number of yes/no questions needed to reduce
uncertainty about the variable's outcomes to zero. The \textit{redundancy} of a
signal denotes the difference how expressive the signal could be (i.e.,
proportional to the number of unique states or symbols used to transmit the
signal) and the actual information rate (i.e., the entropy of each individual
state or symbol). If a brain network's nodes are fully independent, then the
number of bits required to express a single activity pattern is proportional to
the number of nodes. The network would also be minimally redundant, since the
status of every node would be needed to fully express a single brain activity
pattern. If a brain network's nodes are fully coupled and identical, then the
number of bits required to express a single activity pattern is proportional to
the number of unique states or values any individual node can take on. Such a
network would be highly redundant, since knowing any individual node's state
would be sufficient to recover the full-brain activity pattern.

%   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
%     *also* very rich-- e.g., the maximum number of informative components is large
% compression and task demands -- MackEtal20, ZhouEtal19
% superEEG -- prediction of held-out data; implies redundancy
% timecorr -- higher order network patterns during higher order cognition implies more complexity
% cognitive richness and scrambling-- various Hasson lab papers
